---
title: Partisan Bias
author: Angela Zhao
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)
library(haven)
library(lubridate)

# Parameters

file_raw_zipcodes_fire_smoke_survey <- 
  here::here("data/zipcodes_fire_smoke_survey/zipcodes_fire_smoke_survey.shp")

file_zipcodes_ca_geometry <-
  here::here("data/ca_zipcodes/ca_zipcodes.shp")

file_raw_smoke <- here::here("data-raw/mzip_month_smoke.RDS")

CA_ALBERS <- 
  "+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

CA_ALBERS_EPSG <- 3310

survey_rename <-
  c(
    "zipcode_radius (miles)" = 'zpcd_..', 
    num_fires_intersect = 'nm_frs_', 
    zipcode_geometry = 'geometry', 
    wildfire_disturb = 'wldfr_d', 
    module_1_2_order = 'md_1_2_', 
    wildfire_treat = 'wldfr_t', 
    time_in_state = 'tm_n_st', 
    consider_relocate_1 = 'cnsd__1',
    consider_relocate_2 = 'cnsd__2',
    consider_relocate_3 = 'cnsd__3',
    consider_relocate_4 = 'cnsd__4',
    state_relocate = 'stt_rlc', 
    country_relocate_reason_1 = 'cntry_r',
    country_relocate_reason_2 = 'cntry__',
    relocate_pick_most = 'rlct_pck_m', 
    relocate_pick_least = 'rlct_pck_l', 
    parent_guardian_1 = 'prnt__1', 
    parent_guardian_2 = 'prnt__2', 
    parent_guardian_3 = 'prnt__3', 
    parent_guardian_4 = 'prnt__4', 
    grandparent_1 = 'grndp_1',
    grandparent_2 = 'grndp_2',
    grandparent_3 = 'grndp_3',
    grandparent_4 = 'grndp_4',
    own_vs_rent = 'own_vs_', 
    faminc_new = 'fmnc_nw', 
    presvote16post = 'prsvt16', 
    state_lookup = 'stt_lkp', 
    religpew = 'religpw', 
    pew_churatd = 'pw_chrt',                  
    pew_bornagain = 'pw_brng', 
    pew_religimp = 'pw_rlgm', 
    pew_prayer = 'pw_pryr', 
    starttime = 'starttm'
  )

selected_recode_key <-
  c(
   "not selected" = "0", 
    "selected" = "1"
  )

ca_zipcodes <-
  file_zipcodes_ca_geometry %>% 
  read_sf() %>% 
  st_drop_geometry() %>% 
  pull(name)


```

## Reading in and wrangling the data

Columns stt_lk_m, stt_lk_l had suspec values, and also had values that were not contained within the codebook. They were removed. 

```{r}

zipcodes_fire_smoke_survey <-
  file_raw_zipcodes_fire_smoke_survey %>% 
  read_sf() %>% 
  select(-c(stt_lk_m, stt_lk_l)) %>%
  rename(!!! survey_rename) %>% 
  rename_at(
    vars(q1_1_1_1:q12_8_4), 
    ~ str_replace(., "(q\\d*)(_)([:alpha:]*)(.*)", "\\1\\2mod\\4")
  ) %>% 
  mutate(
    zipcode_geometry = st_transform(zipcode_geometry, CA_ALBERS)
  )

st_crs(zipcodes_fire_smoke_survey$zipcode_geometry) <- CA_ALBERS_EPSG

# Checking the geometry was zipcode geometry
# ca <-
#   boundaries(geography = "state", resolution = "500k", projection = "albers") %>%
#   filter(NAME == "California") %>%
#   st_transform(crs = CA_ALBERS)
# 
# zipcodes_fire_smoke_survey %>%
#   ggplot() +
#   geom_sf(data = ca) +
#   geom_sf()


```

## Wildfire perception

### Looking at wildfire perception among party base

```{r}

wildfires_survey_zipcodes <-
  zipcodes_fire_smoke_survey %>% 
  select(
    zipcode:weight, 
    wildfire_in_state = q1_mod1_2_1, 
    personal_experience_wildfire = q1_mod1_2_2, 
    no_experience_wildfire = q1_mod1_2_4, 
    party = pid3, 
    party_strength = pid7
  ) %>% 
  mutate_at(
    vars(wildfire_in_state:no_experience_wildfire), 
    ~ recode(., !!! selected_recode_key) %>% as.double()
  ) %>% 
  filter(no_experience_wildfire == 0, zipcode %in% ca_zipcodes) 

wildfires_survey_zipcodes %>% 
  st_drop_geometry() %>% 
  count(party, wildfire_in_state, wt = weight) %>% 
  group_by(party) %>% 
  mutate(
    prop = n / sum(n)
  ) %>% 
  ungroup() %>% 
  mutate(
    order = if_else(wildfire_in_state == 0, prop, 0), 
    party = fct_reorder(party, order)
  ) %>% 
  filter(wildfire_in_state == 0) %>% 
  ggplot(aes(party, prop)) +
  geom_hline(aes(yintercept = mean(prop)), color = "red") +
  geom_col() +
  labs(
    x = "Political party", 
    y = "Prop of resp, by party, who believed no major wildfires occured", 
    title = 
      "By party, hightest prop of independents believe no major wildfires 2018-2019"
  )

```

### Wildfire perception controlling for people who have personally experienced a fire

```{r}

wildfires_survey_zipcodes %>% 
  st_drop_geometry() %>% 
  filter(personal_experience_wildfire == 1) %>% 
  count(party, wildfire_in_state, wt = weight) %>% 
  group_by(party) %>% 
  mutate(
    prop = n / sum(n)
  ) %>% 
  ungroup() %>% 
  mutate(
    order = if_else(wildfire_in_state == 0, prop, 0), 
    party = fct_reorder(party, order)
  ) %>% 
  filter(wildfire_in_state == 0) %>% 
  ggplot(aes(party, prop)) +
  geom_hline(aes(yintercept = mean(prop)), color = "red") +
  geom_col() +
  labs(
    x = "Political party", 
    y = "Prop of resp, by party, who believed no major wildfires occured", 
    title = 
      "By party, hightest prop of independents believe no major wildfires 2018-2019"
  )


```

## Looking at wildfire perception while controlling for smoke density


```{r}
wildfires_survey_zipcodes %>% 
  st_drop_geometry() %>% 
  filter(
    density != 0, 
    num_fires_intersect != 0,
    party %in% c("Democrat", "Republican", "Independent")
  ) %>% 
  count(`zipcode_radius (miles)`, party, wildfire_in_state, wt = weight) %>% 
  group_by(`zipcode_radius (miles)`, party) %>% 
  mutate(
    prop = n / sum(n)
  ) %>%
  ungroup() %>% 
  mutate(
    order = if_else(wildfire_in_state == 0, prop, 0), 
    party = fct_reorder(party, order)
  ) %>% 
  filter(wildfire_in_state == 0) %>% 
  ggplot(aes(party, prop)) +
  facet_wrap(vars(`zipcode_radius (miles)`)) +
  geom_hline(aes(yintercept = mean(prop)), color = "red") +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(
    x = "Political party", 
    y = "Prop of resp, by party, who believed no major wildfires occured", 
    title = 
      "By party, prop of resp who believed no major wildfires 2018-2019"
  )

wildfires_survey_zipcodes %>% 
  st_drop_geometry() %>% 
  filter(
    density != 0, 
    num_fires_intersect != 0,
    party %in% c("Democrat", "Republican", "Independent")
  ) %>% 
  count(party, wt = weight)


```
```{r}
file_smoke <-
  file_raw_smoke %>% 
  readRDS() %>% 
  mutate(
    date = make_date(year = as.character(year), month = month, day = 01), 
    years = "2018-2019"
  ) %>% 
  rename(
    zipcode = mzip
  ) %>% 
  group_by(years, zipcode, density) %>%
  mutate(
    count = sum(count), 
    days = sum(days)
  ) %>% 
  select(years, zipcode, density, count, days) %>%
  distinct() %>% 
  arrange(years, zipcode, density)

file_smoke
```



```{r}

wildfires_survey_zipcodes %>% 
  st_drop_geometry() %>% 
  filter(
    density != 0, 
    num_fires_intersect != 0,
    party %in% c("Democrat", "Republican", "Independent")
  ) %>% 
  count(`zipcode_radius (miles)`, party, wildfire_in_state, wt = weight) %>% 
  group_by(`zipcode_radius (miles)`, party) %>% 
  mutate(
    prop = n / sum(n)
  ) %>%
  ungroup() %>% 
  mutate(
    order = if_else(wildfire_in_state == 0, prop, 0), 
    party = fct_reorder(party, order)
  ) %>% 
  filter(wildfire_in_state == 0) %>% 
  ggplot(aes(party, prop)) +
  facet_wrap(vars(`zipcode_radius (miles)`)) +
  geom_hline(aes(yintercept = mean(prop)), color = "red") +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(
    x = "Political party", 
    y = "Prop of resp, by party, who believed no major wildfires occured", 
    title = 
      "By party, prop of resp who believed no major wildfires 2018-2019"
  )
```

